---
title: "Step 1. Converting data to hyperframes"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{1_data_to_hfr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style type = "text/css">
h1.title {
  font-size: 30px;
}
h1 { /* Header 1 */
  font-size: 26px;
}
h2 { /* Header 2 */
    font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 16px;
}
div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0; 
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(readr)

load("~/Dropbox/Research/geocausal/Data/airstr.csv")
load("~/Dropbox/Research/geocausal/Data/activ.csv")
load("~/Dropbox/Research/geocausal/Data/iraq_window.Rda")

activ <- activ[, c("latitude", "longitude", "date", "type_out")]
airstr <- airstr[, c("latitude", "longitude", "date", "Type")]
```

```{r setup, include = FALSE}
devtools::load_all()
#library(geocausal)
```

The purpose of this vignette is letting users convert dataframes with spatiotemporal dimensions to hyperframes with point process data. By the end of this vignette, users will be able to do the following tasks:

- task 1
- task 2

The functions that are introduced in this vignette are as follows:

- <tt>get_hfr</tt>

# Loading data

We will use two csv files: <tt>activ.csv</tt> and <tt>airstr.csv</tt>. 

- <tt>activ.csv</tt>: spatiotemporal information of outcomes (i.e., insurgencies) 
- <tt>airstr.csv</tt>: spatiotemporal information of treatment (i.e., airstrikes)

Let's first check what these datasets look like.

## Outcome data

In the dataset <tt>activ</tt>, locations of insurgencies are summarized by their latitudes and longitudes, with their onset captured by dates.

```{r}
head(activ)
```

There are three subtypes of insurgencies (column <tt>type_out</tt>): 

- improvised explosive devices (IED),
- small arms fire (SAF), and 
- other outcomes. 

```{r}
table(activ$type_out)
```

## Treatment data

The treatment data is similar, except for the fact that subtypes are now captured by column <tt>Type</tt>.

```{r}
head(airstr)
```

There are two sub-categories of airstrikes: 

- airstrikes and 
- shows of force (SOF: simulated airstrikes where no weapons are released).

```{r}
table(airstr$Type)
```

# Converting the data to hyperframes

Our first task is to convert data to a collection of point processes. Intuitively, we want to generate maps of all the locations of airstrikes and insurgencies for all subtypes and all dates. For dates and subtypes with no observations, we want to get an empty map.

[Figure here]

To do this, we use <tt>get_hfr</tt> function.

```{r, warning = FALSE}
# Treatment
treatment_hfr <- get_hfr(data = airstr,
                         subtype_column = "Type",
                         window = iraq_window,
                         time_column = "date",
                         time_range = c("2007-02-23", "2008-07-05"),
                         coordinates = c("longitude", "latitude"),
                         combined = TRUE)
```
