---
title: "Step 4. Getting histories"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{4_getting_histories}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style type = "text/css">
h1.title {
  font-size: 30px;
}
h1 { /* Header 1 */
  font-size: 26px;
}
h2 { /* Header 2 */
    font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 16px;
}
div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0; 
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)

library(readr)
library(ggplot2)
library(furrr)
library(spatstat)
```

```{r setup, include = FALSE}
devtools::load_all()
#library(geocausal)
```

```{r load data, include = FALSE}
dat_hfr <- readRDS("~/geocausal/Data/dat_hfr_3.rds")

load("~/geocausal/Data/iraq_window.Rda")

ps_covs <- c("logPopulation", "prior_rivers", "prior_roads", "All_Cities",
             "aid", "prior_PrevKinetic.1", "prior_PrevKinetic.7", "prior_PrevKinetic.30",
             "prior_PrevAttacks.1", "prior_PrevAttacks.7", "prior_PrevAttacks.30",
             "prior_PrevSOF.1", "prior_PrevSOF.7", "prior_PrevSOF.30",
             "time.1", "time.2", "time.3",
             "Settle.IQ.G01", "Settle.IQ.G02", "Settle.IQ.G03", "Settle.IQ.G04",
             "Settle.IQ.G05", "Settle.IQ.G06", "Settle.IQ.G07", "Settle.IQ.G08",
             "Settle.IQ.G09", "Settle.IQ.G10", "Settle.IQ.G11", "Settle.IQ.G12",
             "Settle.IQ.G13", "Settle.IQ.G14", "Settle.IQ.G15", "Settle.IQ.G16",
             "Settle.IQ.G17", "Settle.IQ.G18")
```

As a final step of data preparation, we obtain histories of treatment and outcomes.

# History of treatment and outcomes

We consider the history of treatment and outcomes as covariates. To create columns for the treatment and outcome history, we aggregate treatment and outcomes over 1, 7, and 30 days. To do so, we use <tt>get_history</tt> function by specifying data to convert, the lag, and the window. 

For example, if you want to convert all outcomes (saved as <tt>dat_hfr\$all_outcome</tt>) aggregated over the past 30 days in Iraq, then you set <tt>Xt = dat_hfr\$all_outcome</tt>, <tt>lag = 30</tt>, and <tt>window = iraq_window</tt> (the window is saved as <tt>iraq_window</tt> in our case).

```{r, warning = FALSE}
lags <- c(1, 7, 30)
time_interval <- 1:nrow(dat_hfr)

# History of all outcomes (allout = Activity in the original code)

## Point process
hist_allout_1 <- lapply(time_interval, get_history,
                        Xt = dat_hfr$all_outcome,
                        lag = lags[1], window = iraq_window)
hist_allout_7 <- lapply(time_interval, get_history,
                        Xt = dat_hfr$all_outcome,
                        lag = lags[2], window = iraq_window)
hist_allout_30 <- lapply(time_interval, get_history,
                         Xt = dat_hfr$all_outcome,
                         lag = lags[3], window = iraq_window)

dat_hfr$hist_allout_1 <- purrr::map(hist_allout_1, 1)
dat_hfr$hist_allout_7 <- purrr::map(hist_allout_7, 1)
dat_hfr$hist_allout_30 <- purrr::map(hist_allout_30, 1)

rm(hist_allout_1, hist_allout_7, hist_allout_30)

## Smoothing
dat_hfr$hist_allout_1 <- get_smoothed_outcome(data_interest = dat_hfr$hist_allout_1,
                                              method = "mclust", initialization = TRUE,
                                              sampling = 0.05)
dat_hfr$hist_allout_7 <- get_smoothed_outcome(data_interest = dat_hfr$hist_allout_7,
                                              method = "mclust", initialization = TRUE,
                                              sampling = 0.05)
dat_hfr$hist_allout_30 <- get_smoothed_outcome(data_interest = dat_hfr$hist_allout_30,
                                               method = "mclust", initialization = TRUE,
                                               sampling = 0.05)
```

```{r, warning = FALSE}
# History of airstrikes

## Point process
hist_kinetic_1  <- lapply(time_interval, get_history,
                          Xt = dat_hfr$Airstrike,
                          lag = lags[1], window = iraq_window)
hist_kinetic_7  <- lapply(time_interval, get_history,
                          Xt = dat_hfr$Airstrike,
                          lag = lags[2], window = iraq_window)
hist_kinetic_30  <- lapply(time_interval, get_history,
                           Xt = dat_hfr$Airstrike,
                           lag = lags[3], window = iraq_window)

dat_hfr$hist_kinetic_1 <- purrr::map(hist_kinetic_1, 1)
dat_hfr$hist_kinetic_7 <- purrr::map(hist_kinetic_7, 1)
dat_hfr$hist_kinetic_30 <- purrr::map(hist_kinetic_30, 1)

rm(hist_kinetic_1, hist_kinetic_7, hist_kinetic_30)

## Smoothing
dat_hfr$hist_kinetic_1 <- get_smoothed_outcome(data_interest = dat_hfr$hist_kinetic_1,
                                               method = "mclust", initialization = TRUE,
                                               sampling = 0.05)
dat_hfr$hist_kinetic_7 <- get_smoothed_outcome(data_interest = dat_hfr$hist_kinetic_7,
                                               method = "mclust", initialization = TRUE,
                                               sampling = 0.05)
#dat_hfr$hist_kinetic_30 <- get_smoothed_outcome(data_interest = dat_hfr$hist_kinetic_30,
#                                                method = "mclust", initialization = TRUE,
#                                                sampling = 0.05)
## Returns an error "cannot allocate vector of size 38.6 Gb" but the memory size is infty
```

```{r, warning = FALSE}
# History of SOFs

## Point process
hist_SOF_1  <- lapply(time_interval, get_history,
                      Xt = dat_hfr$SOF,
                      lag = lags[1], window = iraq_window)
hist_SOF_7  <- lapply(time_interval, get_history,
                      Xt = dat_hfr$SOF,
                      lag = lags[2], window = iraq_window)
hist_SOF_30  <- lapply(time_interval, get_history,
                       Xt = dat_hfr$SOF,
                       lag = lags[3], window = iraq_window)

dat_hfr$hist_SOF_1 <- purrr::map(hist_SOF_1, 1)
dat_hfr$hist_SOF_7 <- purrr::map(hist_SOF_7, 1)
dat_hfr$hist_SOF_30 <- purrr::map(hist_SOF_30, 1)

rm(hist_SOF_1, hist_SOF_7, hist_SOF_30)

## Smoothing
dat_hfr$hist_SOF_1 <- get_smoothed_outcome(data_interest = dat_hfr$hist_SOF_1,
                                           method = "mclust", initialization = TRUE,
                                           sampling = 0.05)
dat_hfr$hist_SOF_7 <- get_smoothed_outcome(data_interest = dat_hfr$hist_SOF_7,
                                           method = "mclust", initialization = TRUE,
                                           sampling = 0.05)
dat_hfr$hist_SOF_30 <- get_smoothed_outcome(data_interest = dat_hfr$hist_SOF_30,
                                            method = "mclust", initialization = TRUE,
                                            sampling = 0.05)
```
Now the hyperframe contains all the necessary data to estimate the causal effects of treatment.

```{r}
head(dat_hfr)
```

```{r, include = FALSE}
saveRDS(dat_hfr, file = "~/geocausal/Data/dat_hfr_4.rds")
```