---
title: "Step 6. Obtaining counterfactual densities"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{6_counterfactual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style type = "text/css">
h1.title {
  font-size: 30px;
}
h1 { /* Header 1 */
  font-size: 26px;
}
h2 { /* Header 2 */
    font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 16px;
}
div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0; 
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)

library(readr)
library(ggplot2)
library(furrr)
library(spatstat)
library(lubridate)
library(stars)
library(ggthemes)
library(ggpubr)
```

```{r setup, include = FALSE}
devtools::load_all()
#library(geocausal)
```

```{r load data, include = FALSE}
dat_hfr <- readRDS("~/geocausal/Data/dat_hfr_4.rds")
airstr_base <- get(load("~/geocausal/Data/airstr_base.csv"))
load("~/geocausal/Data/covariates/routes_dist.dat")
load("~/geocausal/Data/iraq_window.Rda")
airstr_base <- airstr_base[, c("latitude", "longitude", "date", "Type")] 
##Data used here is not included in the analysis, also it excludes SOF
```

# 1 Obtaining counterfactual densities

## 1.1 Obtaining the baseline density

- Use the data from 2006 to obtain the baseline density
- Check: Only one density is obtained

```{r, out.height=400, out.width=600, fig.dim = c(9, 6)}
# Data
head(airstr_base)
range(airstr_base$date)

# Get baseline density
baseline <- get_baseline_density(data = airstr_base, 
                                 coordinates = c("longitude", "latitude"),
                                 window = iraq_window)

baseline$density_plot
baseline$density_plot + scale_fill_distiller(type = "seq", direction = -1, palette = "Greys")
baseline$point_plot
```
# 1.2 Counterfactual (1) Changing the expected number of outcomes

First, get the summary and plot so that we can see the expected number of airstrikes.

```{r}
# Get the target number
baseline_summary <- get_baseline_summary(data = airstr_base, 
                                         time_column = "date")

baseline_summary$quantile
baseline_summary$mean
baseline_summary$variance
baseline_summary$plot
```

Then first try the simple multiplication. Since the baseline density is standardized, the target_number is already the target expectation.

```{r}
# Type 1 intervention (simple multiplication)
baseline_density <- baseline$density
target_number <- 3

counterfactual_multiply <- get_counterfactual_density(baseline_density = baseline_density,
                                                      expected_number = target_number,
                                                      window = iraq_window)
counterfactual_multiply$density
counterfactual_multiply$density_plot
counterfactual_multiply$density_plot + scale_fill_distiller(type = "seq", direction = -1, palette = "Greys")
```

Then type-2 interventions.

```{r}
# Type 2 intervention (shifting the location)

## Data on locations
focus_locations <- cbind(longitude = 44.025, latitude = 32.614) #Baghdad only
focus_locations <- cbind(longitude = 43.158, latitude = 36.349) #Mosul only
focus_locations <- cbind(longitude = c(44.025, 43.158), latitude = c(32.614, 36.349)) #Baghdad and Mosul

## Distance from focus
distance_from_focus <- get_focus(focus_locations = focus_locations, 
                                 window = iraq_window)
plot(distance_from_focus)

### Another distance map
distance_from_routes <- exp(-routes_dist)/integral(routes_dist)
plot(distance_from_routes)

target_densities <- list(distance_from_focus, distance_from_routes)
priorities <- c(3, 5)

power_density <- get_power_density(target_densities = target_densities, 
                                   priorities = priorities, 
                                   window = iraq_window)

power_density$density
power_density$density_plot

counterfactual_location <- get_counterfactual_density(baseline_density = baseline_density,
                                                      expected_number = target_number,
                                                      power_density = power_density$density,
                                                      window = iraq_window)

counterfactual_location$density
counterfactual_location$density_plot
```

Simulation and visualization exercises for power density
- Suppose that we keep the distance from cities as it is with alpha = 2 and want to see how changing alpha from 1 to 6 for routes affects the power density...

```{r}
sim_power <- simulate_power_density(target_densities = list(distance_from_focus), #This must be a list element
                                    density_to_manipulate = distance_from_routes,
                                    priorities = 2,
                                    priorities_for_manipulation = c(1, 2, 5, 10, 15, 20),
                                    window = iraq_window)

sim_power
```

Examples
