---
title: "Step 6. Obtaining counterfactual densities"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{6_counterfactual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style type = "text/css">
h1.title {
  font-size: 30px;
}
h1 { /* Header 1 */
  font-size: 26px;
}
h2 { /* Header 2 */
    font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 16px;
}
div#TOC li {
    list-style:none;
    background-image:none;
    background-repeat:none;
    background-position:0; 
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)

library(readr)
library(ggplot2)
library(furrr)
library(spatstat)
library(lubridate)
```

```{r setup, include = FALSE}
devtools::load_all()
#library(geocausal)
```

```{r load data, include = FALSE}
airstr_base <- get(load("~/geocausal/Data/airstr_base.csv"))
load("~/geocausal/Data/iraq_window.Rda")
airstr_base <- airstr_base[, c("latitude", "longitude", "date", "Type")] 
##Data used here is not included in the analysis, also it excludes SOF
```

# 1 Obtaining counterfactual densities


## 1.1 Obtaining the baseline density

- Use the data from 2006 to obtain the baseline density
- Check: Only one density is obtained

```{r, out.height=400, out.width=600, fig.dim = c(9, 6)}
# Data
head(airstr_base)
unique(airstr_base$Type)
range(airstr_base$date)

# Get baseline density
baseline_density <- get_baseline_density(airstr_base, 
                                         coordinates = c("longitude", "latitude"),
                                         window = iraq_window,
                                         return_ppp = FALSE)

plot(baseline_density)
```
# 1.2 Counterfactual (1) Changing the expected number of outcomes




# 1.3 Counterfactual (2) Shifting the geographical focus




# 1.4 Shifting the focus



# 1.5 Everything at once





